\subsection{Flexibility competition and synergy}\label{sec:discussion_RQ1}
\textbf{Research question 1:} \textit{Are there competitive or synergistic effects between different types of flexibility when different technologies provide them?}


The effects of flexibility integration on \ac{PS}, \ac{CS}, \ac{CR} and \ac{SEW} are shown in Table \ref{table:economic_effects}. Interaction between flexibility options and the resulting synergistic and competitive effects are shown in Table \ref{table:synergy_competition}. This table shows whether competition or synergy was observed by the cell colour. The corresponding cell label specifies the affected economic indicator. 
\begin{table}
\centering
\small
\begin{tabular}{llllllll}
\hline
\begin{tabular}[c]{@{}l@{}}\textbf{Flexibility}\\\textbf{option}\end{tabular}              & \begin{tabular}[c]{@{}l@{}}\textbf{Thermal}\\\textbf{PS}\end{tabular}    & \begin{tabular}[c]{@{}l@{}}\textbf{RESE}\\\textbf{PS}\end{tabular} & \begin{tabular}[c]{@{}l@{}}\textbf{H2}\\\textbf{PS}\end{tabular} & \begin{tabular}[c]{@{}l@{}}\textbf{Storage}\\\textbf{PS}\end{tabular} & \textbf{CS}   & \textbf{CR}                       & \textbf{SEW}                      \\ \hline
storage                                                                   
& \cellcolor[HTML]{f27a7d}   
& \cellcolor[HTML]{c5f9d7}  
& \cellcolor[HTML]{C0C0C0}  
& \cellcolor[HTML]{f27a7d} 
& \cellcolor[HTML]{f27a7d} 
& \cellcolor[HTML]{f7d486} 
& \cellcolor[HTML]{c5f9d7} \\ \hline
transmission                                                              & \cellcolor[HTML]{f27a7d}                                & \cellcolor[HTML]{c5f9d7}                          & \cellcolor[HTML]{C0C0C0}                              & \cellcolor[HTML]{c5f9d7}                             & \cellcolor[HTML]{f7d486} & \cellcolor[HTML]{f27a7d} & \cellcolor[HTML]{c5f9d7} \\ \hline
\begin{tabular}[c]{@{}l@{}}storage\\\& transmission\end{tabular}        & \cellcolor[HTML]{f27a7d}                                & \cellcolor[HTML]{c5f9d7}                          & \cellcolor[HTML]{C0C0C0}                              & \cellcolor[HTML]{c5f9d7}                             & \cellcolor[HTML]{c5f9d7} & \cellcolor[HTML]{f27a7d} & \cellcolor[HTML]{c5f9d7} \\ \hline
demand-side                                                               & \cellcolor[HTML]{f27a7d}                                & \cellcolor[HTML]{f27a7d}                          & \cellcolor[HTML]{C0C0C0}                              & \cellcolor[HTML]{f27a7d}                             & \cellcolor[HTML]{c5f9d7} & \cellcolor[HTML]{C0C0C0} & \cellcolor[HTML]{c5f9d7} \\ \hline
\begin{tabular}[c]{@{}l@{}}demand-side\\\& storage\end{tabular}         & \cellcolor[HTML]{f27a7d}                                & \cellcolor[HTML]{c5f9d7}                          & \cellcolor[HTML]{C0C0C0}                              & \cellcolor[HTML]{f27a7d}                             & \cellcolor[HTML]{f27a7d} & \cellcolor[HTML]{f7d486} & \cellcolor[HTML]{c5f9d7} \\ \hline
sector coupling                                                           & \cellcolor[HTML]{c5f9d7}                                & \cellcolor[HTML]{c5f9d7}                          & \cellcolor[HTML]{c5f9d7}                              & \cellcolor[HTML]{f27a7d}                             & \cellcolor[HTML]{f27a7d} & \cellcolor[HTML]{f27a7d} & \cellcolor[HTML]{c5f9d7} \\ \hline
\begin{tabular}[c]{@{}l@{}}sector coupling\\\& storage\end{tabular}    & \cellcolor[HTML]{c5f9d7}                                & \cellcolor[HTML]{c5f9d7}                          & \cellcolor[HTML]{c5f9d7}                              & \cellcolor[HTML]{f27a7d}                             & \cellcolor[HTML]{f27a7d} & \cellcolor[HTML]{f27a7d} & \cellcolor[HTML]{c5f9d7} \\ \hline
\begin{tabular}[c]{@{}l@{}}sector coupling\\\& demand-side\end{tabular} & \cellcolor[HTML]{c5f9d7}                                & \cellcolor[HTML]{c5f9d7}                          & \cellcolor[HTML]{c5f9d7}                              & \cellcolor[HTML]{f27a7d}                             & \cellcolor[HTML]{f27a7d} & \cellcolor[HTML]{f27a7d} & \cellcolor[HTML]{c5f9d7} \\ \hline
    &  &      &      &     &    &    &           \\
\cellcolor[HTML]{f27a7d}{\color[HTML]{000000} Decrease}                   
& \cellcolor[HTML]{c5f9d7}{\color[HTML]{000000} Increase} 
& \multicolumn{3}{l}{\cellcolor[HTML]{f7d486}{\color[HTML]{000000} Location-dependent}}  
& \multicolumn{3}{l}{\cellcolor[HTML]{C0C0C0}{\color[HTML]{000000} Not specified}}                   \\ \hline
\end{tabular}
\caption{Economic effects of flexibility integration}
\label{table:economic_effects}
\end{table}

\begin{table}
\centering
\small
\begin{tabular}{lllll}
\hline
\begin{tabular}[c]{@{}l@{}}\textbf{Flexibility}\\\textbf{option}\end{tabular} 
& \textbf{Transmission}  
& \textbf{Storage} 
& \textbf{Demand-side}                             
& \begin{tabular}[c]{@{}l@{}}\textbf{Sector}\\\textbf{coupling}\end{tabular}
\\ \hline
transmission  
& \cellcolor[HTML]{C0C0C0}  
& \cellcolor[HTML]{f7d486}{\color[HTML]{000000} SEW} 
& \cellcolor[HTML]{C0C0C0}  
& \cellcolor[HTML]{C0C0C0}   \\ \hline
storage   
& \cellcolor[HTML]{f7d486}{\color[HTML]{000000} SEW} 
& \cellcolor[HTML]{C0C0C0}       
& \cellcolor[HTML]{c5f9d7}{\color[HTML]{000000} CS} 
& \cellcolor[HTML]{f27a7d}{\color[HTML]{000000} SEW}     \\ \hline
demand-side                                                  
& \cellcolor[HTML]{C0C0C0}                           
& \cellcolor[HTML]{c5f9d7}{\color[HTML]{000000} CS}  
& \cellcolor[HTML]{C0C0C0}                                          
& \cellcolor[HTML]{c5f9d7}{\color[HTML]{000000} RESE PS} \\ \hline
sector coupling                                              
& \cellcolor[HTML]{C0C0C0}                           
& \cellcolor[HTML]{f27a7d}{\color[HTML]{000000} SEW} 
& \cellcolor[HTML]{c5f9d7}{\color[HTML]{000000} RESE PS}            
& \cellcolor[HTML]{C0C0C0}                               \\ \hline
  &    &      &      &    \\
\cellcolor[HTML]{f27a7d}{\color[HTML]{000000} Competition}   
& \cellcolor[HTML]{c5f9d7}{\color[HTML]{000000} Synergy}                    
& \multicolumn{2}{l}{\cellcolor[HTML]{f7d486}{\color[HTML]{000000} Location-dependent}} 
& \cellcolor[HTML]{C0C0C0}{\color[HTML]{000000} Not specified}                  \\ \hline
\end{tabular}
\caption{Synergy and competition of flexibility integration}
\label{table:synergy_competition}
\end{table}
\FloatBarrier
The results indicate that additional storage decreases \ac{CS}, thermal \ac{PS} and storage \ac{PS} while \ac{RESE} \ac{PS} rises. The latter benefits from curtailment reductions and fewer extremely low electricity prices. Reducing peak prices lowers thermal power plant \ac{PS}. These effects are a disadvantage for storage, as their arbitrage potential decreases, indicating that additional storage increases overall \ac{SEW}, but competition between existing and new storage always emerges. Moreover, the storage location heavily influences the \ac{SEW} increase related to the annual price duration curve. Storage \ac{PS} reduces as other flexibilities are used. Demand-side flexibility and sector coupling reduce price volatility and consequent storage \ac{PS}. 


The location-dependent difference in \ac{SEW} increase of 6 M\euro{}  is reduced to 1 M\euro{} by expanding the interconnector. Enhanced trading capacity leads to a price convergence between bidding zones; therefore, a completely non-congested electricity market would elicit location-independent results. Consequently, storage with a congested transmission line generates different welfare gains based on the bidding zone. Hence, synergistic effects can occur between storage and transmission line expansion.


Demand-side flexibility particularly benefits end-users while reducing \ac{PS}. End-users' ability to shift demand to periods with lower electricity prices has a significant potential for them. While the overall \ac{SEW} increases by 2 M\euro{} , \ac{CS} increases by 8 M\euro{} . In combination with storage, further positive synergistic effects for end-users occur. While storage expansion reduces \ac{CS} by 32 M\euro{}, the simultaneous implementation of storage expansion and demand-side flexibility reduces \ac{CS} by 19 M\euro{}. Therefore, \ac{CS} is 13 M\euro{} higher through demand-side flexibility, indicating that the negative effects of producer-side flexibility, which lower \ac{CS}, can be significantly reduced by demand-side flexibility. Combining demand-side flexibility with hydrogen sector coupling does not further increase \ac{CS}. These flexibility options compete, as hydrogen production is another application of demand-side flexibility. Overall \ac{SEW} increases show neither competitive nor synergistic effects in every combination of different flexibility types. 


Hydrogen sector coupling leads to massive \ac{SEW} gains. The main beneficiaries are thermal power plants, \ac{RESE}, and electrolysers. Electrolysers generate a \ac{PS} on the hydrogen market through hydrogen production, which occurs when electricity prices would be low and increases prices, leading to drastically higher thermal and \ac{RESE} \ac{PS}. As additional electricity demand shrinks price volatility, storage \ac{PS} reduces, generating competition between storage and sector coupling. Moreover, the hydrogen production in each bidding zone leads to a price convergence that reduces electricity trade and consequently the \ac{CR}. 


Transmission line expansion only leads to minor changes in overall \ac{SEW} combined with hydrogen sector coupling. The additional electricity demand to produce hydrogen increases the benefits of transmission line expansion for end-users. \ac{CS} reduction lowers from 3 513 M\euro{}  to 2 622 M\euro{}  as a result of enhanced trading capacity. Sole trading capacity expansion leads only to a 48 M\euro{} \ac{CS} increase. The overall \ac{SEW} increase is slightly influenced by transmission line expansion as \ac{PS} decreases to compensate for the \ac{CS} increase.


If applied comprehensively, combining storage expansion and demand-side flexibility can replace transmission line expansion and vice versa. While storage and demand-side flexibility increase \ac{SEW} by 32 M\euro{}  without transmission line expansion, the same flexibility increases \ac{SEW} by only 17 M\euro{}  if the interconnector is expanded. The price convergence generated by the implementation of sector coupling has neither synergistic nor competitive effects on the \ac{SEW} through transmission line expansion. 
\FloatBarrier